
    This is based off of Aaron's framework
    
    Introduction:
        Genetic variation causes
        Genetic variation between individuals and coalescent theory
        Impact of genetic variation within the genome: a chromosomal translocation event and mixed lineage leukemia.
            - 

    Part 1: Interpreting genetic variation between individuals 
        Chapter 1: A particle filter for demographic inference
        Chapter 2: Ancient admixture 
    Part 2: The impacts of genetic variation within a single genome
        Chapter 3: LDA for bulk ATAC-seq and Erythropoesis
        Chapter 4: LDA for MLL-r 
        Chapter 5: A NN for blah
    Part 3: End matter

    For the MLL part:
        Genomic regulation by chromatin conformation and accessible DNA.
            How is it regulated? What are the effects?
            Measuring it via ATAC-seq and DNA-ase seq
        Topic modelling (I can save this for the chapter)
            General idea
            LDA 
            Adjusting the algorithm for bulk ATAC-seq. (This is due to a lack of appropriate differentiation systems for single cell. )
        MLL leukemia biology and pathology (This can be )
            Also will need to introduce erythropoesis too if this will be a big part of it.
            Will this be a seperate chapter? Or a part of the topic modelling chapter?

        * My big problem is that this isn't really related to genetic variation.... It's more
            like finding regions associated with a cell type. 

