\subsection*{Personal}

I owe a massive debt of gratitude to my supervisors Professors Gerton Lunter and Thomas Milne for their persistent support and encouragement. Gerton has been both an invaluable source of knowledge and a staunch optimist throughout the many challenges of this degree. His guidance through adversity has taught me to be a better scientist and a more mature individual. Tom became an unexpected and invaluable mentor in the second half of my DPhil. Endlessly generous with his intricate knowledge of cancer biology, Tom's unique combination of kindness, enthusiasm, and generosity helped me to discover my own passion for research. Special thanks to Professor Jim Hughes who took on an unofficial advisory role. His knowledge of both machine learning and functional genomics, as well as his sense of humour, laid the foundation for an extremely beneficial collaboration. 

My fellow labmates have been crucial to any success that I've had throughout this degree. Perhaps none more so than Alastair Smith, whose creativity and mentorship extensively contributed to our work on MLL-AF4 leukemia. Joe (Sha) Zhu has always been generous with his time and knowledge, and along with Donna Henderson laid the groundwork for the SMC2 method and the applications in this thesis. The Lunter group has been my home away from home, with special thanks to Richard Brown, Ron Schwessinger, and Daniel Cooke for the endless hours of table tennis at the Wellcome Centre. 

I was lucky enough to form many lasting friendships in Oxford, particularly in the Genomic Medicine and Statistics cohort with Xilin, Dan, Chris, and Ryan. The pandemic spread us across the globe but they always found time to offer help, advice, or a joke. Thank you also to my friends in Canada, who found time to suffer through my endless and changing passions in genetics. My family and soon to be in-laws in Ottawa and beyond have provided me with both a solid support structure and the opportunity to explore my interests overseas. 

However, none of the above would have been in any way possible without the unending support and patience of my fiancée Kennedy (Ning) Hao. She has been both my rock and my best friend for the last decade, and I hope to be able to provide the same support to her.  


\subsection*{Institutional}
I am incredibly grateful to my funders and institutional supporters for the opportunity to pursue this research. Firstly to the Wellcome Trust and the Clarendon Scholarship for allowing me to attend the University of Oxford and removing any barriers to pursuing my interests. Secondly to the Maple Leaf Fund and the Victor Dahdaleh Foundation for their support through the Canadian Centennial Scholarship Fund. Thirdly to the Wellcome Centre for Human Genetics and the Weatherall Institute of Molecular Medicine, with particular thanks to the Centre for Computational Biology and the staff involved in maintaining their compute cluster as well as the biomedical research computing (BMRC) system.